PInSoRo: A Dataset for Deep-learning of Human-Robot Social Interactions

### 2017-06-27 paper.tex
Polishing

In 2015, the Google DeepMind team demonstrated how a convolutional
recurrent neural network could learn to play the game Break-Out (amongst
48 other Atari games) by only \emph{looking} at the gaming console
screen~\cite{mnih2015human}. This result represents a major milestone: they show
that a relatively small sample size (about 500 games) is sufficient for an 
artificial agent to not only learn how to play (which requires an implicit model 
of time to adequately move the Break-Out paddle), but to also create gaming 
strategies that \emph{look like} they would necessitate planning (the system first
breaks bricks on one side to eventually get the ball to break-out and reach the area
\emph{above} the remaining bricks, therefore ensuring rapid progress in the
game).

More recently, Ogata's team~\cite{yang2017repeatable} has demonstrated how an
adequately configured RNN is able to learn a complex robotic task (folding soft
objects like towels using a dual-arm mobile manipulator) from only \emph{35}
demonstrations, one demonstration being a $\approx$ 70 seconds-long teleoperated
sequence. The network inputs are the raw video stream from the head camera and the
12 DoF of the two arms and the total training dataset comprises of about 28000
steps of data. Successfully folding towels entails an explicit sequencing of
actions (therefore implicit temporal modeling). The fact that such a complex
process can be successfully learned from a very small dataset certainly opens
many prospects.

Indeed, we argue that the complexity of mechanisms that such neural networks
have been able to uncover and model should invite our community to question its
applicability to human-robot interactions in general, and sustained, natural
child-robot interactions in particular.  However, the current lack of an
appropriate HRI dataset suitable for the training of neural networks hinders
this investigation. We present hereafter such a dataset, and detail our
acquisition methodology.

@@

The use of interaction datasets to teach robots how to socially behave has been
previously explored, and can be considered as the extension of the traditional
learning from demonstration (LfD) paradigms to social interactions (for
instance~\cite{nehaniv2007imitation,mohammad2015interaction}). Previous examples
have generally focused on low-level recognition or generation of short,
self-standing behaviours, including social gestures~\cite{nagai2005learning} and
gazing behaviours~\cite{calinon2006teaching}.

Based on a human-human interaction dataset, Liu \etal~\cite{liu2014how} have
investigated machine learning approaches to learn longer interaction sequences.
Using unsupervised learning, they train a robot to act as a shop-keeper,
generating both speech and socially acceptable motions. Their approach remains
task-specific, and while they report only limited success, they emphasise the
``life-likeness'' of the generated behaviours.

Kim \etal~\cite{kim2015pororobot} highlight that applying deep learning to
visual scene information in an HRI scenario was successful, but that
generating behaviours for the robot to be able to act in a dynamic and uncertain
environment remains a challenge.
They rely on unsupervised learning, with a
training stage consisting in clustering behavioural states (a state consist in
\emph{behavioural elements} (joint states of the two interacting agents,
associated to the current speech elements, as well as the current joint
\emph{spatial formation}), followed by the \todo[inline]{finish that}.

@@

Using the {\tt gazr} face tracking library\footnote{Available online from
\url{https://github.com/severin-lemaignan/gazr}. {\tt gazr} relies on the {\tt
dlib} face tracker for facial features
extraction.}~\cite{lemaignan2016realtime}, children faces and facial features can
are extracted and reprojected onto the 3D point-clouds (Fig.~\ref{fig|rviz},
right). Faces have been successfully detected in 57\% of the recorded frames
in the dataset. Undetected faces are due to occlusions (the child put her arm in
front of the camera), large head rotations or quick child motion.

>>>>

In 2015, the Google DeepMind team demonstrated how a convolutional
recurrent neural network could learn to play the game Break-Out (amongst
48 other Atari games) by only looking at the gaming console
screen~\cite{mnih2015human}. This result represents a major milestone: they show
that a relatively small sample size (about 500 games) is sufficient for an 
artificial agent to not only learn how to play (which requires an implicit model 
of time to adequately move the Break-Out paddle), but to also create gaming 
strategies that look like they would necessitate planning (the system first
breaks bricks on one side to eventually get the ball to break-out and reach out the area
above the remaining bricks, therefore ensuring rapid progress in the
game).

More recently, Ogata's team~\cite{yang2017repeatable} has demonstrated how an
adequately configured RNN is able to learn a complex robotic task (folding soft
objects like towels using a dual-arm mobile manipulator) from only \emph{35}
demonstrations of $\approx$ 70 seconds-long teleoperated
sequences. The network inputs are the raw video stream from the head camera and the
12 DoF of the two arms. Successfully folding towels entails an explicit sequencing of
actions (therefore implicit temporal modeling): the fact that such a complex
process can be successfully learned from a small training dataset should lead
us to reconsider the range of domains to which RNNs would be applicable.

Indeed, as we have previously argued~\cite{lemaignan2016towards}, we believe
that the complexity of mechanisms that such neural networks have been able to
uncover and model should invite our community to question its applicability to
human-robot interactions in general, and sustained, natural child-robot
interactions in particular. The current lack of appropriate HRI
datasets suitable for neural networks training is however an obstacle: we
introduce in this article the PInSoRo dataset, a large dataset of child-child
natural interactions suitable for machine learning.

@@

The use of interaction datasets to teach robots how to socially behave has been
previously explored, and can be considered as the extension of the traditional
learning from demonstration (LfD) paradigms to social interactions (for
instance~\cite{nehaniv2007imitation,mohammad2015interaction}). Existing
research focuses however on low-level recognition or generation of short,
self-standing behaviours, including social gestures~\cite{nagai2005learning}
and gazing behaviours~\cite{calinon2006teaching}.

Based on a human-human interaction dataset, Liu \etal~\cite{liu2014how} have
investigated machine learning approaches to learn longer interaction sequences.
Using unsupervised learning, they train a robot to act as a shop-keeper,
generating both speech and socially acceptable motions. Their approach remains
task-specific, and while they report only limited success, they emphasise the
``life-likeness'' of the generated behaviours.
%They rely on unsupervised learning, with a
%training stage consisting in clustering behavioural states (a state consist in
%\emph{behavioural elements} (joint states of the two interacting agents,
%associated to the current speech elements, as well as the current joint
%\emph{spatial formation}), followed by the

@@

Using the {\tt gazr} face tracking library\footnote{Available online from
\url{https://github.com/severin-lemaignan/gazr}. {\tt gazr} relies on the {\tt
dlib} face tracker for facial features
extraction.}~\cite{lemaignan2016realtime}, children faces and facial features can
are extracted and reprojected onto the 3D point-clouds (Fig.~\ref{fig|rviz},
right). Faces have been successfully detected in 56\% of the recorded frames
in the dataset. Undetected faces are due to occlusions (the child put her arm in
front of the camera), large head rotations or quick child motion.
